Measuring Individual Partisan Harm in Redistricting

### 2021-09-11 paper/paper.Rmd
add intro; join refs

What is the question?

Districting is a complicated process.
Residents of a state or locality must be partitioned into districts.
These districts must adhere to many constraints, often requiring components such as contiguity and equal population.
In drawing the districts, it's easy to lose sight of the individuals impacted by district drawing.
Often, we consider how parties succeed or fail under the electoral environments generated by enacted plans.
Yet this misses that parties are formed by voters who ally with parties.
Rather than considering parties, we ask how individuals are harmed or benefited by the districting process?

Why does it matter?

Why is it hard?
Decades of

What did we do?

What did we learn?

What is the paper's contribution?

>>>>

Districting is a complicated process.
Residents of a state or locality must be partitioned into districts based on the geographies that they live in.
These districts must adhere to many constraints, often requiring components such as contiguity and equal population.
In drawing the districts, it's easy to lose sight of the individuals impacted by district drawing.
Often, we consider how parties succeed or fail under the electoral environments generated by enacted plans.
Yet this misses that parties are formed by voters who ally with parties.
Rather than considering parties, we ask how individuals are harmed or benefited by the districting process?

Decades of literature on redistricting have focused on the fairness to parties.
This literature defers to symmetric outcomes for parties, commonly referred to as partisan symmetry \cite{a:bunch:of:king}.
Partisan symmetry is a useful idea which greatly improves academic understanding of the redistricting process.
We Why is it hard?
Understanding the redistricting process requires a feasible and reliable understanding of the counterfactual set of redistricting plans.
Early methodologies for simulating plans relied on randomness to generate many different plans and enhanced understanding of redistricting greatly \\citep{ciri:etal:00, chen:rodd:13, defO:duch:solo:21.
These methods greatly enhanced the understanding of redistricting within contexts of a particular state with set populations.
Yet, without a known target distribution, these heuristic approaches to simulation fall do not offer necessary guarantees for confidence in the simulated plans to represent the total space of possible plans \cite{fife:etal:20b}.
Recent advances in Markov Chain methods as applied to redistricting allow us move beyond simulation towards sampling plans from this definable space \\cite{fife:etal:20, mcca:imai:20, autr:etal:21}.

Using sampling methods, we can then develop a reliable counterfactual that can be tuned to consider the samping constraints imposed by the laws and practices of a state.
This counterfactual allows us to characterize the likely plans from a state.
Often, we can would then call this an ensemble analysis, which is a flexible approach to sampling redistricting plans by simulation, relying on large numbers of plans for stability \cite{matt:etal:21} or sufficient diversity of nearly-independent plans \cite{kenn:etal:forth}.
using this counterfactual, we thus create measures of harm, utility, and fairness which are based on a single plan compared to a set of other reasonable and feasible plans which encorporate the necessary legal constraints to make a plan viable.

We offer a reminder that geographic context is key to understanding common problems in political science..
Elections themselves are geographically determined.
Recent research offers important resolution of the degree of partisan sorting in the US \cite{brow:enos:something}.
By combining those understandings with the necessities of redistricting, we are able to expand the localized aspect of harm.
Further, we suggest that while the efficiency gap may not incorporate geographic information, it often helps capture individual harm better than other common measures.
This is a great benefit as an indicator to detect extreme gerrymanders in cases where sufficiently local geographic data to simulate plans is not available.\
We further clarify that common existing measures of partisan fairness -- while they may capture the difficulties offered to a party -- do not necessarily capture the representational harms incurred by individuals who live in districts which do not at all represent their preferences.

Our metrics offer a framework for understanding approximately individual harms, benefits, and net fairness.
Using counterfactual plans to understand this space allows us to properly calibrate understandings of the impact of redistricting plans on individual understanding.
Thus, we create measures which capture the possibility that no plans exist which favor one community \cite{duch:massachusetts}.
Using this class of metrics which describe the individual affect via representation allow us to better align legalistic interpretations of harm with political science understandings.
